---
title: "Tutorial4_Spatial regression analysis"
author: "Zeyu(Zey)"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 4.1 Constructing a Spatial Weight Matrix

```{r}

library(spdep)
library(spatialreg)

#read data (it is already in the program so do not need to download it)
data(columbus)
mydata <- columbus

#use "attach" to prevent typing "$" for every variables
attach(mydata)

#Col.gal.nb provides an adjacency list of neighbors. We give it a better name
neighbors <- col.gal.nb

#check the class of the adjacency list
class(neighbors)


```

Note: it is an ‘nb’ or ‘neighborhood’ object. Usually we need to first convert a dataframe to ‘nb’, and then to a ‘weighted list - listw’ object.

```{r}
#dataframe to nb

#class(dataframe) <- “nb”
#nb2listw(nblist)
```

Another approach (which is more suitable when having isolates) is to first make an adjacency matrix. Then you can use: listw\<-mat2listw(adjacencymatrix).

```{r}
#convert nb to listw
#style = "W" means row-standardize
listw <- nb2listw(neighbors, style = "W")
summary(listw)
```

We see that an average region has about 5 neighbors.

## 4.2 Linear Regression Analysis

```{r}
summary(CRIME)
summary(INC)
summary(HOVAL)

plot(neighbors, coords)

olsreg <- lm(CRIME ~ INC + HOVAL)
summary(olsreg)
```

## 4.3 Spatial Analysis

### 4.3.1 Moran’s I test

```{r}
#see whether we need a spatial model in the first place
moran.test(CRIME, listw)

#null hypothesis: no spatial dependence
# p<0.05  null hypothesis rejected, we should use spatial model
```

### 4.3.2 Lagrange multiplier tests for spatial lag and spatial error dependencies

```{r}
#which type of spatial models we should use
lm.LMtests(olsreg, listw, test=c("LMlag", "LMerr"))

#null hypothesis:no spatial lag dependence
#null hypothesis:no spatial error dependence

#both rejected in this case
```

### 4.3.3 Spatial lag model

```{r}
slm <- lagsarlm(CRIME ~ INC + HOVAL, data = mydata, listw, Durbin = "FALSE")
summary(slm)

#inspect direct and indirect effect
#use a simulation with 200 iterations to obtain p-values
# ‘zstats = TRUE’ to obtain p-values
im<-impacts(slm, listw = listw, zstats=TRUE, R=200)
sums<-summary(im,  zstats=T)

#effect value(simulated)
data.frame(sums$res)

#effect p-value(simulated)
data.frame(sums$pzmat)
```

### 4.3.4 Spatial error model

```{r}
sem <- errorsarlm(CRIME ~ INC + HOVAL, data = mydata, listw, Durbin = FALSE)
summary(sem)
```

### 4.3.5 Spatial autoregressive combined model

```{r}
sac <- sacsarlm(CRIME ~ INC + HOVAL, data = mydata, listw, Durbin = FALSE)
summary(sac)
```

\*both rho and lamda are insignificant now

```{r}
#Impact: distinguish direct and indirect effects
im2<-impacts(sac, listw = listw, zstats=TRUE, R=200)
sums2<-summary(im2,  zstats=T)
data.frame(sums2$res)
data.frame(sums2$pzmat)
```

### 4.3.6 Spatial Durbin model

```{r}
sdm <- lagsarlm(CRIME ~ INC + HOVAL, data = mydata, listw, Durbin = TRUE)
summary(sdm)

```

```{r}
#Impact: distinguish direct and indirect effects
im3<-impacts(sdm, listw = listw, zstats=TRUE, R=200)
sums3<-summary(im3,  zstats=T)
data.frame(sums3$res)
data.frame(sums3$pzmat)
```

### 4.3.7 Spatial Durbin error model

```{r}
sdem <- errorsarlm(CRIME ~ INC + HOVAL, data = mydata, listw, Durbin = TRUE)
summary(sdem)
```

### 4.3.8 General nested model

```{r}
gnm <- sacsarlm(CRIME ~ INC + HOVAL, data = mydata, listw, Durbin = TRUE)
summary(gnm)
```

```{r}
#Impact: distinguish direct and indirect effects
im4<-impacts(gnm, listw = listw, zstats=TRUE, R=200)
sums4<-summary(im4,  zstats=T)
data.frame(sums4$res)
data.frame(sums4$pzmat)
```

## 4.4 Knecht Assignment

1&2.

```{r}
library(igraph)
library(haven) # to read Stata files
library(reshape2)
library(tidyverse)

PupilsWaveV <- read_dta(file = "PupilsWaveV.dta")

grep("acthomew", names(PupilsWaveV), value = TRUE)

table(PupilsWaveV$acthomew)
table(PupilsWaveV$sex)

PupilsWaveV <- subset(PupilsWaveV, acthomew >= 1 & acthomew <= 5)
PupilsWaveV$female <- ifelse(PupilsWaveV$sex == 2, 0, 1)

table(PupilsWaveV$acthomew)
table(PupilsWaveV$female)

PupilsWaveV$acthomew_r <- 6 - PupilsWaveV$acthomew

table(PupilsWaveV$acthomew_r)
table(PupilsWaveV$female)

```

```{r}
attach(PupilsWaveV)

summary(acthomew_r)
summary(female)
summary(age)

reg <- lm(acthomew_r ~ female + age)
summary(reg)
```

3\.

```{r}
library(lme4)

multi <- lmer(formula = acthomew_r ~ 1 + female + age + (1 | schoolnr), 
               data = PupilsWaveV, REML = FALSE)

summary(multi)
```

4\.

```{r}
#filter class 12b
c12b <- subset(PupilsWaveV, schoolnr == "12b")

library(network)
library(igraph)
library(reshape2)
library(tidyverse)


elist <- c12b %>% 
select(namenr, friend1:friend12) %>%   # select only the " best friends"  network
  pivot_longer(c(friend1:friend12)) %>% 
  filter(!is.na(value)) %>% # drop the missings
  rename(from ="namenr", to = "value", sourcevar= "name") %>% #just nice for interpretation
  relocate(to, .after=from) #move around the columns
nodelist <- c12b %>% 
  select(namenr,sex, age) %>% 
  rename(name = "namenr")
classnet <- graph_from_data_frame(elist, vertices = nodelist)

adjmatrix <- as_adjacency_matrix(classnet, sparse = FALSE)
adjmatrix

class(adjmatrix)
```

5\.

```{r}
library(spdep)
library(spatialreg)

listw <- mat2listw(adjmatrix, style = "W", zero.policy = TRUE)

summary(listw)
```

6\.

```{r}
moran.test(c12b$acthomew_r, listw, zero.policy = TRUE)
#no no spatial dependence

reg12b <- lm(acthomew_r ~ female+age, data = c12b)
lm.LMtests(reg12b, listw, test = c("LMlag", "LMerr"), zero.policy = TRUE)
# spatial lag dependence exists
# no spatial error dependence
```

7\.

```{r}
slm <- lagsarlm(acthomew_r ~ female + age, data = c12b, listw, Durbin = "FALSE", zero.policy = TRUE)
summary(slm)

#inspect direct and indirect effect
#use a simulation with 200 iterations to obtain p-values
# ‘zstats = TRUE’ to obtain p-values
im<-impacts(slm, listw = listw, zstats=TRUE, R=200)
sums<-summary(im,  zstats=T)

#effect value(simulated)
data.frame(sums$res)

#effect p-value(simulated)
data.frame(sums$pzmat)
```

8\.

```{r}
sem <- errorsarlm(acthomew_r ~ female + age, data = c12b, listw, Durbin = FALSE, zero.policy = TRUE)
summary(sem)
```

```{r}
sac <- sacsarlm(acthomew_r ~ female + age, data = c12b, listw, Durbin = FALSE, zero.policy = TRUE)
summary(sac)

#Impact: distinguish direct and indirect effects
im<-impacts(sac, listw = listw, zstats=TRUE, R=200)
sums<-summary(im,  zstats=T)
data.frame(sums$res)
data.frame(sums$pzmat)
```

9\.

```{r}
sdm <- lagsarlm(acthomew_r ~ female + age, data = c12b, listw, Durbin = TRUE, zero.policy = TRUE)
summary(sdm)
```

```{r}
sdem <- errorsarlm(acthomew_r ~ female + age, data = c12b, listw, Durbin = TRUE, zero.policy = TRUE)
summary(sdem)
```

```{r}
gnm <- sacsarlm(acthomew_r ~ female + age, data = c12b, listw, Durbin = TRUE, zero.policy = TRUE)
summary(gnm)
```

10. maybe we should not use spatial model for this class 12b data. There is no spatial dependence at all.
